Sample-efficient Reinforcement Learning \\ with Implicitly Quantized Representations

### 2024-02-02 paper/main.tex
udpated

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Sample-Efficient RL with Implicitly Quantized Representations}

@@

\twocolumn[
% \icmltitle{Submission and Formatting Instructions for \\
%            International Conference on Machine Learning (ICML 2024)}
% \icmltitle{Sample-efficient Reinforcement Learning with Latent Representations}
\icmltitle{Sample-Efficient Reinforcement Learning \\ with Implicitly Quantized Representations}

@@

\textbf{Representation learning}
Our representation learning is based solely on the latent-state consistency loss,
%
\begin{align} \label{eq:rep-loss}
  \mathcal{L}_{\text{rep}}(\theta, \phi; \tau)
 % &= \E_{s_{t} \sim \mathcal{D}} \left[ \sum_{h=0}^{H-1} \| \hat{z}_{t+1} - z_{t+1} \|_{2}^{2} \right] \\
%&= \E_{(s_{t}, a_{t}, s_{t+1}) \sim \mathcal{D}}
% &= \E_{\tau \sim \mathcal{D}}
% \left[ \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2} \right]
&= \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2},
\end{align}
%
which minimizes the distance between the next state predicted by the transition model $d_{\phi}(e_{\phi}(z_{t}), a_{t})$
and the next state predicted by the momentum encoder $e_{\bar{\phi}}(z_{t+1})$.
The prediction $d_{\phi}(e_{\phi}(z_{t}), a_{t})$ uses the online encoder which is being trained jointly with the transition model.
The target $e_{\bar{\phi}}(z_{t+1})$ is calculated with the momentum encoder which uses an exponential moving average (EMA)
of the encoder's weights $\bar{\theta} \leftarrow \tau \bar{\theta} + (1-\tau)\theta$.
The target network update rate is denoted $\tau$.
Unlike previous works, we normalize the latent vector

@@

\textbf{Policy and value function learning}

\begin{align} \label{eq:value-loss}
  \mathcal{L}_{Q}(\psi ; \mathcal{D}) &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\theta}(s_{t}), a_{t}) - y)^{2}  \right], \quad \forall k \in 1, 2 \\
  y &= \sum_{n=0}^{N-1} \gamma^{n} r_{t+n} + \gamma^{n} \min_{k \in \{1,2\}} Q_{\bar{\psi}_{k}}(e_{\bar{\theta}}(s_{t+n}), a_{t+n})
\end{align}

\begin{align} \label{eq:policy-loss}
 \mathcal{L}_{\pi}(\eta; \mathcal{D}) = \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\theta}(o_{t}), a_{t}) \right]
\end{align}

>>>>

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Sample-Efficient Reinforcement Learning with Implicitly Quantized Representations}

@@

\twocolumn[
% \icmltitle{Submission and Formatting Instructions for \\
%            International Conference on Machine Learning (ICML 2024)}
% \icmltitle{Sample-efficient Reinforcement Learning with Latent Representations}
\icmltitle{Sample-efficient Reinforcement Learning \\ with Implicitly Quantized Representations}

@@

\textbf{Representation learning}
Our representation learning is based solely on the latent-state consistency loss,
%
\begin{align} \label{eq:rep-loss}
  \mathcal{L}_{\text{rep}}(\theta, \phi; \tau)
 % &= \E_{s_{t} \sim \mathcal{D}} \left[ \sum_{h=0}^{H-1} \| \hat{z}_{t+1} - z_{t+1} \|_{2}^{2} \right] \\
%&= \E_{(s_{t}, a_{t}, s_{t+1}) \sim \mathcal{D}}
% &= \E_{\tau \sim \mathcal{D}}
% \left[ \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2} \right]
&= \sum_{h=0}^{H-1} \gamma^{h} \| d_{\phi}(e_{\theta}(s_{t}), a_{t}) - e_{\bar{\theta}}(s_{t+1}) \|_{2}^{2},
\end{align}
%
which minimizes the distance between the next state predicted by the transition model $d_{\phi}(e_{\phi}(z_{t}), a_{t})$
and the next state predicted by the momentum encoder $e_{\bar{\phi}}(z_{t+1})$.
The prediction $d_{\phi}(e_{\phi}(z_{t}), a_{t})$ uses the online encoder which is being trained jointly with the transition model.
The target $e_{\bar{\phi}}(z_{t+1})$ is calculated with the momentum encoder which uses an exponential moving average (EMA)
of the encoder's weights $\bar{\theta} \leftarrow \tau \bar{\theta} + (1-\tau)\theta$.
The target network update rate is denoted $\tau$.

Unlike previous works, we normalize the latent vector

@@

\textbf{Policy and value function learning}
We learn the policy and value function using TD3 \citep{fujimotoAddressingFunctionApproximation2018}.
However, we follow \citet{yaratsMasteringVisualContinuous2021,zhaoSimplifiedTemporalConsistency2023}
and augment the loss with n-step returns.
The only difference to TD3 is that instead of using the original observations $o_{t}$, we map them through the
encoder $z_{t} = e_{\bar{\theta}}(o_{t})$ and learn the value function and policy in the latent space $z_{t}$.
The value function is updated by minimizing the following objective:
\begin{align} \label{eq:value-loss}
  \mathcal{L}_{Q} &= \E_{\tau \sim \mathcal{D}} \left[ (Q_{\psi_{k}}(e_{\bar{\theta}}(s_{t}), a_{t}) - y)^{2}  \right], \quad  \forall k \in 1, 2 \\
  y &= \sum_{n=0}^{N-1} \gamma^{n} r_{t+n} + \gamma^{n} \min_{k \in \{1,2\}} Q_{\bar{\psi}_{k}}(e_{\bar{\theta}}(s_{t+n}), a_{t+n}), \nonumber
\end{align}
where we use policy smoothing by adding clipped Gaussian noise $\epsilon \sim \mathcal{N} (0,\sigma^{2})$ to the
action $a_{t+n} = \pi_{\eta}(z_{t+n}) + \epsilon$.
Note that we use the target encoder to get the latent states in both the prediction and the target.

\begin{align} \label{eq:policy-loss}
 \mathcal{L}_{\pi} = \E_{(o_{t}, a_{t}) \sim \mathcal{D}} \left[ \min_{k\in\{1,2\}} Q_{\psi_{k}}(e_{\theta}(o_{t}), a_{t}) \right]
\end{align}
